---
title: "Example 2"
author: "Sachi"
date: "7/31/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generate double decrement table

### Independent rates

Need a vector of ages $x$ and the two vectors of independent rates, $q_x^d$ and $q_x^w$. The $q_x^d$ are given by Australian insured 2004--8 ultimate (2+) death only non-smoker rates for males. The $q_x^w$ are given by $$q_{[x]+t}^w = \max{(0.05, 0.2-0.02t)},\qquad t=0,1,\ldots,9.$$

```{r}
x <- c()
x[50:60] <- 50:60
x

qx.d.ind <- c()

qx.d.ind[50] <- 0.00142
qx.d.ind[51] <- 0.00156
qx.d.ind[52] <- 0.00172
qx.d.ind[53] <- 0.00189
qx.d.ind[54] <- 0.00208
qx.d.ind[55] <- 0.00230
qx.d.ind[56] <- 0.00255
qx.d.ind[57] <- 0.00286
qx.d.ind[58] <- 0.00320
qx.d.ind[59] <- 0.00358
qx.d.ind

qx.w.ind <- c()

qx.w.ind[50:59] <- pmax(.05,0.2-.02*0:9)
qx.w.ind

```

### Dependent rates

We can now calculate the multiple decrement table. first for deaths.

```{r}
qx.d.dep <- c()
qx.d.dep[50:59] <- qx.d.ind[50:59]*(1-0.5*qx.w.ind[50:59])
qx.d.dep
```
And then for withdrawals
```{r}
qx.w.dep <- c()
qx.w.dep[50:59] <- qx.w.ind[50:59]*(1-0.5*qx.d.ind[50:59])
qx.w.dep
```
## Double decrement table
```{r}
al.x <- c()
al.x[50] <- 1000000
```
initialise $(ad)_x^d$ and $(ad)_x^w$
```{r}
ad.x.d <- c()
ad.x.w <- c()

for (j in 50:60) {
  ad.x.d[j] <- al.x[j]*qx.d.dep[j]
  ad.x.w[j] <- al.x[j]*qx.w.dep[j]
  al.x[j+1] <- al.x[j]-ad.x.d[j]-ad.x.w[j]
}
al.x
al.x[50:60]
```

this is probably a good time to pretty print the results
```{r}
# pretty
```
### Calculating the required policy values

foo
```{r}
qx <- c()

qx[50] <- 0.00478880
qx[51] <- 0.00537740
qx[52] <- 0.00603064
qx[53] <- 0.00675456
qx[54] <- 0.00755572
qx[55] <- 0.00844128
qx[56] <- 0.00941902
qx[57] <- 0.01049742
qx[58] <- 0.01168566
qx[59] <- 0.01299373

qx
```
and we need the life table corresponding to these $q_x$ values.
```{r}
lx <- c()

lx[50] <- 100000

dx <- c()

for (j in 50:60) {
  dx[j] <- lx[j]*qx[j]
  lx[j+1] <- lx[j]-dx[j]
}
lx
```
to calculate policy values and surrender values we need the annuity-due term life annuity values on these two bases.
```{r}
a.due.PV <- c()
a.due.SV <- c()

px <- c()
px[50:60] <- 1-qx[50:60]
px
```
here are the life annuity due values on the policy value basis (i.e., $i$=4.25\%).
```{r}
a.due.PV[59] <- 1
v <- 1/(1.0425)

for (j in 59:51) {
  a.due.PV[j-1] <- 1+v*px[j-1]*a.due.PV[j]
}
a.due.PV

a.due.SV[59] <- 1
v <- 1/1.09

for (j in 59:51) {
  a.due.SV[j-1] <- 1+v*px[j-1]*a.due.SV[j]
}
a.due.SV[60] <- 0
a.due.SV
```
good! now we're in a position to do all our cash flow analysis!

## In-force cash flows

For this ten year endowment policy, we need the vectors of

1. Premia
2. Expenses
3. Death benefits
4. Withdrawal benefits
5. Maturity benefits
6. Policy values

So, here goes. First, vector of premia
```{r}
prems.BoY <- c(rep(7600,10))
prems.BoY
```
and expenses
```{r}
expenses.BoY <- c()
expenses.BoY[1] <- 0.7*prems.BoY[1]+20
expense.inflation <- 1.07^(1:9)
expense.inflation
expenses.BoY[2:10] <- 20*expense.inflation
expenses.BoY
```
and death benefits
```{r}
death.EoY <- c(rep(100000,10))
death.EoY
```
and maturity benefits
```{r}
maturity.EoY <- c(rep(0,10))
maturity.EoY[10] <- 100000
maturity.EoY
```
now we can do the complicated ones, the withdrawl benefit and the policy value. First, the withdrawal benefit, which is given on the surrender value basis ($i=$9%), using a net premium approach
$$S\!V=S(1-\ddot{a}_{x+t\,:\,\overline{n-t}|}/\ddot{a}_{x\,:\,\overline{n}|})$$
where $x=50$, $n=10$ and $t=1,\ldots,10$.
```{r}
wdrwl.EoY <- c()
wdrwl.EoY[1] <- 0
wdrwl.EoY[2:10] <- death.EoY[2:10]*(1-a.due.SV[52:60]/a.due.SV[50])
wdrwl.EoY
```
and for the policy value we have
```{r}
policy.EoY <- death.EoY[1:9]*(1-a.due.PV[51:59]/a.due.PV[50])
policy.EoY[10] <- 0
policy.EoY
```
So, for an in-force policy, we know all the relevant cash flows. Hence, we're now able to move on to the expected case for a single sold policy.

## Expected amounts per sold policy

Here we need vectors of

1. expected premium
2. expected expenses
3. expected withdrawal benefits
4. expected maturity benefits
5. expected interest earned
6. expected policy value (to set aside)
7. expected transfer (like profit)

To get these expected values, the key will be to have ${}_t(ap)_x$. Lucky we have our double decrement table set up and ready and rearing to go! We nedd ${}_1(ap)_{50}$ to {}_10(ap)_{50}$.
```{r}
ap.x <- al.x[51:60]/al.x[50]
ap.x
```
And so we can calculate the expected premium vector easily as
```{r}
Eprem.BoY <- c()
Eprem.BoY[1] <- prems.BoY[1]
Eprem.BoY[2:10] <- prems.BoY[2:10]*ap.x[1:9]
Eprem.BoY
```
and expected expenses as
```{r}
Eexp.BoY <- c()
Eexp.BoY[1] <- expenses.BoY[1]
Eexp.BoY[2:10] <- expenses.BoY[2:10]*ap.x[1:9]
Eexp.BoY
```
The expected maturity benefit paid also depends on survival of the purchaser of the single sold policy.
```{r}
Emat.EoY <- maturity.EoY*ap.x[1:10]
Emat.EoY
```

Expected death benefits paid on one sold policy will require probabilities of death from our cohort of fifty-year olds ${}_t|(aq)^d_{50}$, for $t=0,\ldots,9$.
```{r}
aq.x.d <- ad.x.d[50:59]/al.x[50:59]
Edeath.EoY <- death.EoY[1:10]*ad.x.d[50:59]/al.x[50]
Edeath.EoY 
```
Expected withdrawal benefits paid on one sold policy will work similarly.

```{r}
Ewdrwl.EoY <- c()
Ewdrwl.EoY <- wdrwl.EoY[1:10]*ad.x.w[50:59]/al.x[50]
Ewdrwl.EoY
```
The funds generated by the single policy sold also have to set up reserves: that is, the policy value. This also depends on survivorship. So the expected policy values created from a single sold policy will also be generated by ${}_t(ap)_x$.
```{r}
Epolval.EoY <- policy.EoY*ap.x
Epolval.EoY
```

The assets (net, at beginning of year) from this single sold policy will earn interest every year (at the earned rate). But that's only the annual cash inflow. There's also the policy value that's sitting around that's earning interest, too, for years 2 and beyond.
```{r}
i <- 0.08

Einterest <- c()
Einterest[1] <- (Eprem.BoY[1]-Eexp.BoY[1])*i
Einterest[2:10] <- (Eprem.BoY[2:10]-Eexp.BoY[2:10]+Epolval.EoY[1:9])*i
Einterest
```
So, now we can work out how much money we expect to make on this single sold policy. This last element is the expected transfer vector.
```{r}
EPolValChg <- c()
EPolValChg[1] <- Epolval.EoY[1]
EPolValChg[2:10] <- Epolval.EoY[2:10]-Epolval.EoY[1:9]
EPolValChg
ETransfer <-Eprem.BoY-Eexp.BoY-Edeath.EoY-Ewdrwl.EoY-Emat.EoY+Einterest-EPolValChg
ETransfer
```
## Measures of profitability

### Profit margin
given by the present value of the profit signature divided by the PV of expected premiums.
```{r}
EPVPremsER <- sum(Eprem.BoY/(1+i)^(0:9))
EPVPremsER
EPVPremsOth <- sum(Eprem.BoY/(1+0.1)^(0:9))
EPVPremsOth
EPVTransfersER <- sum(ETransfer/(1+i)^(1:10))
EPVTransfersOth <- sum(ETransfer/(1+0.1)^(1:10))
EPVTransfersER
EPVTransfersOth
ProfMargER <- EPVTransfersER/EPVPremsER
ProfMargER
ProfMarOth <- EPVTransfersOth/EPVPremsOth
ProfMarOth
```

### NPV of Transfers
```{r}
EPVTransfersER <- sum(ETransfer/(1+i)^(1:10))
EPVTransfersOth <- sum(ETransfer/(1+0.1)^(1:10))
EPVTransfersER
EPVTransfersOth
```

### Yield on transfers
```{r}
yield <- 0.1274
EPVTransfers <- ETransfer/(1+yield)^(1:10)
EPVTransfers
EPVoTransfers <- sum(EPVTransfers)
EPVoTransfers
```
